### 2017-11-13 supp.tex
17th

\documentclass[10pt, letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.
\usepackage{bm}
% Include other packages here, before hyperref.
% zizhao package
\usepackage{tabularx} % in the preamble
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}
\usepackage{array,booktabs,arydshln,xcolor} % for widening table line
\usepackage{amsthm}
\usepackage{epstopdf}
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{multicol}
\usepackage[algo2e]{algorithm2e}
\usepackage[toc,page]{appendix}
\usepackage{blindtext}
% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{2871} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{1}
\begin{document}
%\tableofcontents
%\blinddocument

%%%%%%%%% TITLE
%\title{Supplementary Materials for "Understanding, Describing, and Visualizing Medical Images Through Reading Diagnostic Report Text"}

%\author{First Author\\
%Institution1\\
%Institution1 address\\
%{\tt\small firstauthor@i1.org}
%% For a paper whose authors are all at the same institution,
%% omit the following lines up until the closing ``}''.
%% Additional authors and addresses can be added with ``\and'',
%% just like the second author.
%% To save space, use either the email address or home page, not both
%\and
%Second Author\\
%Institution2\\
%First line of institution2 address\\
%{\tt\small secondauthor@i2.org}
%}

%\maketitle
%\thispagestyle{empty}

%appendicesappendices%%%%%%%% ABSTRACT
%\begin{abstract}
%   The ABSTRACT is to be in fully-justified italicized text, at the top
%   of the left-hand column, below the author and affiliation
%   information. Use the word ``Abstract'' as the title, in 12-point
%   Times, boldface type, centered relative to the column, initially
%   capitalized. The abstract is to be in 10-point, single-spaced type.
%   Leave two blank lines after the Abstract, then begin the main text.
%   Look at previous CVPR abstracts to get a feel for style and length.
%\end{abstract}

%%%%%%%%% BODY TEXT
\begin{appendices}
%\renewcommand{\thechapter}{A\arabic{chapter}}
\section{Results on the CUB bird dataset}

\end{appendices}

%{\small
%	\bibliographystyle{ieee}
%	\bibliography{egbib}
%}

\end{document}