### 2017-11-15 cvpr2018.tex
almostfinal

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.
\usepackage{tabularx} % in the preamble
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}
\usepackage{array,booktabs,arydshln,xcolor} % for widening table line
\usepackage{amsthm,bm}
\usepackage{epstopdf}
\usepackage[flushleft]{threeparttable}
\usepackage{amsthm}
\usepackage{multirow}

\usepackage{float}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{capt-of}% or \usepackage{caption}
\usepackage{booktabs}
\usepackage{varwidth}
\usepackage{wrapfig}
\usepackage{makecell}

\usepackage{algpseudocode,algorithm,algorithmicx}
\newcommand*\DNA{\textsc{dna}}

\newcommand*\Let[2]{\State #1 $\gets$ #2}
\algrenewcommand\algorithmicrequire{\textbf{Precondition:}}
\algrenewcommand\algorithmicensure{\textbf{Postcondition:}}

\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

\usepackage{bm}

\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

%\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{2823} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
%\setcounter{page}{4321}
\begin{document}

%%%%%%%%% TITLE
\title{Photographic Text-to-Image Synthesis \\ with a Hierarchically-nested Adversarial Network}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
This paper presents a novel method to deal with the challenging task of generating photographic images conditioned on semantic image descriptions.
Our method introduces accompanying hierarchical-nested adversarial objectives inside the network hierarchies, which regularize mid-level representations and assist generator training to capture the complex image statistics. We present an extensile single-stream generator architecture to better adapt the jointed discriminators and push generated images up to high resolutions. We adopt a multi-purpose adversarial loss to encourage more effective image and text information usage in order to improve the semantic consistency and image fidelity simultaneously. Furthermore, we introduce a new visual-semantic similarity measure to evaluate the semantic consistency of generated images, which also alleviates the cost of human evaluation. With extensive experimental validation on three public datasets, our method significantly improves previous state of the arts on all datasets over different evaluation metrics.

\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
Photographic text-to-image synthesis is a significant problem in generative model research \cite{reed2016generative}, which aims to learn a mapping from a semantic text space to a complex RGB image space. This task requires the generated images to be not only realistic but also \textit{semantically consistent}, For example, the generated images should preserve the object shape and color information described in the text.

%the generated images preserve the object color and sketch described in text and fine-grained descriptive details in image pixels as well.

%However, this problem still remains unsolved due to its particular challenges. 
%insufficient methods are developed to successfully address this task due to its particular challenges.

%It is a special case of generative learning tasks, which aim to map a low-dimensional embedding to a complex RGB image space. Differently, the embedding space of the former is induced by semantic descriptions. Therefore, this task requires the generated images to be \textit{semantically consistent}, i.e., the generated images not only preserve sketch concepts of descriptions but also fine-grained descriptive details in image pixels.

\begin{figure}[t]
%	\begin{center}
	\centering
	\begin{subfigure}[t]{0.5\textwidth}
	%	\centering
		\includegraphics[width=0.95\textwidth]{figure/intro2.pdf}
	\end{subfigure} \\
\centering
	\begin{subfigure}[t]{0.5\textwidth}
	%	\centering
		\includegraphics[width=0.95\textwidth,height=0.55\textwidth]{figure/single_view.pdf}
	\end{subfigure}
	\vspace{-.6cm}
	\caption{Top: Overview of our hierarchically-nested adversarial network. Bottom: Two generated results where fine-grained details are highlighted. \label{fig:intro}} 	
	\vspace{-.5cm}
	%\end{center}
\end{figure}

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=0.4\textwidth]{figure/intro.pdf}
%	\vspace{-.3cm}
%	\caption{Overview of our hierarchically-nested adversarial network. It contains a single-stream generator and several nested discriminators along the depth dimension of the generator, which can generate high-resolution images conditioned on a sentence.} \label{fig:archs-review} \vspace{-.3cm}
%\end{figure}
%\begin{figure}[t]
%	\centering
%	\includegraphics[width=0.4\textwidth]{figure/single_view.pdf}
%	\vspace{-.3cm}
%	\caption{Overview of our hierarchically-nested adversarial network. It contains a single-stream generator and several nested discriminators along the depth dimension of the generator, which can generate high-resolution images conditioned on a sentence.} \label{fig:archs-review} \vspace{-.3cm}
%\end{figure}
Recently, Generative adversarial networks (GANs) have become the main solution to this task. 
Reed \etal \cite{reed2016generative} address this task through a GAN based framework. But this method only scales up to $64^2$ resolution and can barely generate vivid object details.
Based on this method, StackGAN  \cite{han2017stackgan} propose a two-stage training strategy to generate high-resolution $256^2$ images. But this methods requires training two separate GANs.  Later on, \cite{dong2017semantic} proposes to bypass the difficulty of learning mappings from vector embeddings to RGB images and treat it as a pixel-to-pixel translation problem \cite{isola2016image}. It works by re-rendering an arbitrary-style $128^2$ training image conditioned on a targeting description. However, its high-resolution synthesis capability remains unclear.

This paper pays attention to two major difficulties for text-to-image synthesis with GANs. The first is balancing the convergence between generators and discriminators \cite{goodfellow2014generative,improvedGAN}, which is actually a long-term problem in GANs. The second is stably modeling the huge pixel space in high-resolution images and guaranteeing semantic consistency \cite{han2017stackgan}. 
An effective strategy to regularize generators is critical to stabilize the training and help capture the complex image statistics \cite{huang2016stacked}.

In this paper, we propose a novel end-to-end method that can directly model high-resolution image statistics and generate semantically consistent and photographic images (see Figure \ref{fig:intro}). The contributions are described as follows.

Our generator resembles a simple vanilla GAN, without without requiring multi-stage training and multiple internal text conditioning like \cite{han2017stackgan} or additional class label supervision like \cite{dash2017tac}. To tackle the problem of the big leap from the text space to the high-resolution image space, our insight is to leverage and regularize hierarchical representations with additional `deep' adversarial constraints. 
We introduce accompanying hierarchically-nested discriminators at multi-scale intermediate layers to play adversarial games and jointly encourage the generator approaching the real training data distribution. 
We also propose a new convolutional neural network (CNN) design for the generator to support accompanying discriminators more effectively in an end-to-end training manner.
To guarantee the image diversity and semantic consistency, we enforce discriminators at multiple side outputs of the generator to simultaneously differentiate real-and-fake image-text pairs as well as real-and-fake local image patches. %For more details, please refer to Section~\ref{sec:localloss}%focal

%%\textcolor{red}{This multi-purpose conditional adversary is mutually beneficial to allow each focus on its respect duty.}%focal ranges
%To alleviate the need for manual human evaluation about the matching between generated images and text description, we propose to learn a visual-semantic embedding model, thus providing a measure for quantitative evaluation of the image sentence consistency.
%% It is a regularization to limit the discriminators to use unexpected information to make decision.

We validate our proposed method on three datasets, CUB birds \cite{welinder2010caltech}, Oxford-102 flowers \cite{Nilsback08}, and large-scale MSCOCO \cite{lin2014microsoft}. In complement of existing evaluation metric (e.g. Inception score \cite{improvedGAN}) for generative models, we also introduce a new visual-semantic similarity metric to evaluate the alignment between generated images and conditioned text. Extensive experimental results and analysis demonstrate the effectiveness of our method and significantly improved performance compared against previous state of the arts on all three evaluation metrics.

All source code will be released.

\begin{figure*}[t]
	\centering
	\includegraphics[width=0.95\textwidth]{figure/views2.pdf}
	\vspace{-.2cm}
	\caption{Overviews of some typical GAN frameworks. \textbf{A} uses multi-stage GANs \cite{han2017stackgan,denton2015deep}. \textbf{B} uses multiple discriminators with one generator \cite{durugkar2016generative,tu_etal_nips17_d2gan}. \textbf{C} progressively trains symmetric discriminators and generators \cite{Karras2017progressive,huang2016stacked}. \textbf{A} and \textbf{C} can be viewed as decomposing high-resolution tasks to multi-stage low-to-high resolution tasks.  \textbf{D} is our proposed framework that uses a single-stream generator with hierarchically-nested discriminators trained end-to-end.} \label{fig:archs-review} \vspace{-.4cm}
\end{figure*}

\section{Related Work}
%We discusse related work and further clarify the novelty of our method by comparing with them.

Deep generative models have attracted wide interests recently, including GANs \cite{goodfellow2014generative}, Variational Auto-encoders (VAE) \cite{kingma2013auto}, etc \cite{oord2016pixel}. 
There are substantial existing methods investigating better usage of GANs for different applications, such as image synthesis \cite{radford2015unsupervised, shrivastava2016learning}, (unpaired) pixel-to-pixel translation \cite{isola2016image,zhu2017unpaired}, medical applications \cite{costa2017towards}, etc \cite{ledig2016photo,huang2016stacked}.

Text-to-image synthesis is an interesting application of GANs. Reed \etal \cite{reed2016generative}  first introduce a method that can generate $64^2$ resolution images, which is similar with DCGAN \cite{radford2015unsupervised}. This method presents a new strategy for image-text matching aware adversarial training. Reed \etal \cite{reed2016learning} propose a generative
adversarial what-where network (GAWWN) to enable location and content instructions in text-to-image synthesis. StackGAN \etal \cite{han2017stackgan} propose a two-stage stacking GAN training approach that is able to generate $256^2$ compelling images. Recently, Dong \etal \cite{dong2017semantic} propose to learn a joint embedding of images and text so as to re-render a prototype image conditioned on a targeting description. Cha \etal \cite{char2017perceptual} explore the usage of the perceptional loss \cite{johnson2016perceptual} with a CNN pretrained on ImageNet  and Dash \etal \cite{dash2017tac} make use of auxiliary classifiers (similar with \cite{odena2016conditional}) to assist GAN training for text-to-image synthesis.

Learning a continuous mapping from a low-dimensional embedding manifold to a complex real data distribution is a long-standing problem. Although GANs have made significant progress, there are still many unsolved difficulties, e.g. training instability and high-resolution generation. Wide methods have been proposed to address the training instability, such as various training techniques \cite{salimans2016improved,arjovsky2017wasserstein,berthelot2017began,shrivastava2016learning,odena2016conditional}, regularization using extra knowledge (e.g. image labels, ImageNet CNNs, etc) \cite{dosovitskiy2016generating,ledig2016photo,dash2017tac,dash2017tac}, or different generator and discriminator combinations  \cite{metz2016unrolled,durugkar2016generative,yang2017lr,huang2016stacked}. \textit{While our method shows a new way to unite generators and discriminators and does not requires any extra knowledge apart from training paired text and images.} In addition, it is easy to see the training difficulty increases significantly as the targeting image resolution increases.

%for example, label smoothing and discriminator feature matching \cite{salimans2016improved, }, training stronger discriminator is beneficial to encourage generators \cite{arjovsky2017wasserstein}. Balancing through a equilibrium enforcing method \cite{berthelot2017began}
%Preventing discriminators from forgetting past samples and re-introduces artifacts \cite{shrivastava2016learning}. Increasing the information richness in GANs has been shown very useful \cite{odena2016conditional}.

To synthesize high-resolution images, cascade networks are effective to decompose originally difficult tasks to multiple subtasks (Figure \ref{fig:archs-review} A).
Denton \etal \cite{denton2015deep} train a cascade of GANs in a Laplacian pyramid framework (LAPGAN) and use each to synthesize and refine image details and push up the output resolution stage-by-stage. StackGAN also shares some similar idea with LAPGAN. Inspired by this strategy, Chen \etal \cite{chen2017photographic} present a cascaded refinement network to synthesize high-resolution scene from semantic maps. 
%%% Huang \etal \cite{huang2016stacked} propose a top-down stacked GAN to leverage mid-level representations, which shares some similarities with StackGAN and our method. However, this method needs multiple symmetric bottom-up pre-trained discriminators and the usage for high-resolution image is unclear. 
Recently, Karras \etal \cite{Karras2017progressive} propose a progressive training of GANs. The idea is to add symmetric generator and discriminator layers gradually for high-resolution image generation (Figure \ref{fig:archs-review} C). \textit{Compared with these strategies that train low-to-high resolution GANs stage-by-stage or progressively, our method has the advantage of leveraging mid-level representations to encourage the integration of multiple subtasks, which makes end-to-end high-resolution image synthesis in a single vanilla-like GAN possible.}

%More advantages of our method compared with others will be demonstrated in the following.

Leveraging hierarchical representations of neural networks is an effective way to enhance implicit multi-scaling and ensembling for tasks such as image recognition \cite{lee2015deeply} and pixel or object classification \cite{xie2015holistically,cai2016unified,long2015fully}. Particularly, using deep supervision \cite{lee2015deeply} in hierarchical convolutional layers provides short error paths and increases the discriminativeness of feature representations. 
\textit{Our hierarchically-nested adversarial objective is inspired by these families of deeply-supervised CNNs. }
%%%Our proposed hierarchically-nested adversarial supervision enhances layer representations to support final high resolution outputs.

%
%
%\subsection{Summarize different architectures}
%refer to HED.

%\section{Preliminary}
%\subsection{Generative Adversary Network}
%Generative adversary networks (GANs) are composed of a generator $G$ and a discriminator $D$, which are alternatively trained to compete with each other in a two-player minimax game. The discriminator $D$ is optimized to distinguish the synthesized images from the real images, meanwhile, the generator $G$ is trained to fool the discriminator by trying to convert an input distribution $p_z$ to the true data distribution $p_{data}$. Concretely, the overall training procedure can be analogous to the following two-player min-max game $V(D, G)$, where the optimal $G$ and $D$ can be obtained by 
%\begin{equation}
%\label{equ:GAN}
%\begin{split}
%G^*, {D}^* = \arg~\underset{G}{\min}\ \underset{\mathcal{D}}{\max}~ V(D, G)
%\end{split}
%\end{equation}
%
%%\begin{equation}
%%\label{game}
%%\begin{split}
%%\underset{G}{\min}\ \underset{D}{\max}\ V(D, G) =  &\mathbb{E}_{x\sim p_{data}}[\log D(x)] +  \\
%%&\mathbb{E}_{z\sim p_{z}}[\log (1-D(G(z)))]		   
%%\end{split}
%%\end{equation}
%%where $x$ is a real image sampled from the true data distribution $p_{data}$, $z$ is a noise vector usually sampled from some predefined distribution $p_{z}$ (e.g., Gaussian distribution).
%%In practice, It's been shown to work better for the generator $G$ to minimizing $-\log(D(G(z)))$ rather than $\log(1-D(G(z)))$, namely, the $\log(D)$ trick.
%%In this paper, we replace the negative log likelihood loss with a least square objective\cite{lsgan}. 
%
%Conditioned GAN \cite{isola2016image} is a variation of traditional GAN where both the generator network $G$ and discriminator network $D$ are conditioned on an additional condition vector besides the noise vector. In this paper, we consider the task of training a deep  generative adversarial network conditioned on the embedding of the text descriptions of the images.

\section{Method}
%Figure \ref{fig:archs} illustrate the overall method.

\subsection{Adversarial objective basics}
In brief, a GAN \cite{goodfellow2014generative} consists of a generator $G$ and a discriminator $D$, which are alternatively trained to compete with each other. $D$ is optimized to distinguish synthesized images from real images, meanwhile, $G$ is trained to fool $D$ by synthesizing fake images. Concretely, the optimal $G$ and $D$ can be obtained by playing the following two-player min-max game,

\begin{equation}
\label{equ:GAN}
\begin{split}
G^*, D^* = \arg~\underset{G}{\min}\ \underset{D}{\max}~ \mathcal{V}(D, G, Y, \bm z),
\end{split}
\end{equation}
where $Y$ and $\bm z \sim \mathcal{N}(0,1)$ denote training images and random noises, respectively. 
%$G$ is learned to approximate $G(\bm z) \sim p(Y)$.  The loss $\mathcal{L}$ for the valuation function is usually a cross-entropy loss, such that $\mathcal{V}(\cdot) = - \mathcal{L}(\cdot)$. 
$\mathcal{V}$ is the overall GAN objective, which usually takes the form $\mathbb{E}_{Y\sim p_{data}}\big[\log D(Y)\big] + \mathbb{E}_{\bm z\sim p_{z}}\big[\log (1-D(G(\bm z)))\big]$ (cross-entropy loss) or other variations \cite{lsgan,berthelot2017began}.

%Conditioned GAN \cite{isola2016image} is a variation of traditional GAN where both the generator network $G$ and discriminator network $D$ are conditioned on an additional condition vector besides the noise vector. In this paper, we consider the task of training a deep generative adversarial network conditioned on the embedding of the text descriptions of the images.

%To generate high-resolution photo-realistic images from text embeddings, we propose a simple and effective Hierarchically-nested Generative Adversarial Network.
%Our approach is to train both generator and discriminator conditioned on sentence embedding vectors. We leverage functional loss for discriminator and hierarchically nested side outputs as hierarchical supervision for both generator and discriminator to achieve our goal of generating photo-realistic images from text descriptions.  
%
%In general, the generator takes text embedding and random noise vector as input and synthesis images at different scales (e.g, $64\times64$, $128\times128$, $256\times256$). The discriminator, on the other hand, is trying to judge the pair of synthesized images and the corresponding text embedding as fake pairs as well as differentiating the real images from the fake images. $G$ and $D$ are optimized alternatively until converges.
%
%
%\textbf{Notation } 
%We denote the training data set by $S=\{({X}^n, {T}^n),n=1,...,N\}$, where $X^n$ is the $n-$th training image, and $T^n={t^1,...,t^m}$ as $m$ text descriptions corresponds to $X^n$. 
%
%Since we obtain the text embedding vector using the per-trained char-rnn text encoder provided in \cite{reed2016generative}. For simplicity, we do not differentiate text description and it's corresponding text embedding vector in this paper.  Our goal is to learn a generator $G$ that could produce high resolution images of high visual quality as well as variability giving an input text description $t$.  
%
%The input to $G$ compress input sentence embedding, and one noise input.
%Denote the image dimension, text embedding dimension and the dimension of the noise input as $I$, $T$ and $Z$, respectively. Please note that, $G$ has multiple branches and generates multiple images of different sizes. For each image size, we have a distinct discriminator $D$ associated with it. 
%%For simplicity, we only describe one image size.  
%
%The generator is denoted $G: \mathcal{R}^{Z}: \times \mathcal{R}^{T}\rightarrow\mathcal{R}^{I}$, the image discriminator $D_{img}: \mathcal{R}^{I}\rightarrow\{0, 1\}$, and the matching-aware pair discriminator $D_{pair}: \mathcal{R}^{I}: \times \mathcal{R}^{T}\rightarrow\{0, 1\}$. 
%
%\textbf{Variational text embedding } Following the practice of \cite{han2017stackgan}, instead of using the plan deterministic non-linear transformation of the original $t$, we obtain the conditional vector $c$ by sampling from a Gaussian distribution $\mathcal{N}(\mu(t), \Sigma(t) )$, in which the mean $\mu(t)$ and the diagonal covariance matrix $\Sigma(t)$ are non-linear transformation of the original text embedding $t$. This type of conditional augmentation technique can be used to inject randomness into the text embedding vector $t$ to mitigate the common problem that conventional conditional GAN training paradigm usually learns to totally ignore the randomness, resulting to collapsing results lacking diversity \cite{reed2016generative}. In addition, using this kind of text embedding makes the input pure stochastic, making it more difficult for the generator to over fit the training data.

\subsection{Hierarchical-nested adversarial objectives}
Numerous GAN methods have demonstrated ways to unite generators and discriminators for image synthesis. Figure \ref{fig:archs-review} and Section 2 discuss some typical frameworks.
Our method actually explores a new dimension of playing this adversarial game along the depth of a generator (Figure \ref{fig:archs-review} D), which integrates additional hierarchical-nested  discriminators at intermediate layers of the generator. 
The proposed objectives act as regularizers to the hidden space of $\mathcal{G}$, which also offer a short path for better error signal flow and helps reduce the training instability.

The proposed $\mathcal{G}$ is a CNN (defined in Section 3.4) producing multiple outputs with gradually increasing size:
%This architecture resemble the proposed inverse architectures in edge detection\cite{xie2015holistically} and classification \cite{lee2015deeply}, which demonstrate that additional supervision to hidden layers can improve model generalization and boost optimization.  
\begin{equation}
\label{side}
X_1,..., X_s = \mathcal{G}(\bm t, \bm z), 
\end{equation}
where  $\bm t\sim p_{data}$ denotes a sentence embedding (generated by a pre-trained char RNN text encoder\cite{reed2016generative}). $\mathcal{G}$ produces $s-1$ side outputs $\{X_1,...,X_{s-1}\}$  and one final output $X_s$ with the highest resolution.

%$\mathcal{X}^i = \{X^i_1, .. X^i_s\}$ are side output
%Given $z\in \mathcal{R}^{Z} \scriptsize{\sim} \mathcal{N}(0,1)$ and a text description embedding $t$, 
%$G$ generates $s$ side outputs using intermediate layers, $o=\{o_1,...,o_s\}$,
%Note that the dimension of $o_1,..,o_s$ increases according to the resolution
%(e.g., 64$\times$64, 128$\times$128, and 256$\times$256). 
%Denote $\mathcal{X}^i = \{X^i_1, .. X^i_s\}$ as the pyramid of resized images constructed from original image $X^i$.  
%Denote $\mathcal{D}=\{D_1, ..., D_o\}$ as the whole set of discriminators.
%e aim to solve the following optimization problem:

For each side output $X_i$ from the generator, a distinct discriminator $D_i$ is associated to compete with it. Therefore, our full min-max objective is defined as 
\begin{equation}
\label{equ:optim}
\begin{split}
  \mathcal{G}^*, \mathcal{D}^*&  =  \arg~\underset{\mathcal{G}}{\min}\ \underset{\mathcal{D}}{\max}~ \mathcal{V}(\mathcal{G},\mathcal{D}, \mathcal{Y}, \bm t, \bm z), \\
  %\text{where} \;\;	 \mathcal{D} & =  \{D_1, ..., D_s\}, \; \mathcal{Y} = \{Y_1, ..., Y_s\},
\end{split}
\end{equation}
where $\mathcal{D}  =  \{D_1, ..., D_s\} $ and $\; \mathcal{Y} = \{Y_1, ..., Y_s\}$ denotes one training image at multi-scales, $\{1,...,s\}$, forming an image pyramid.
Compared with Eq. (\ref{equ:GAN}), our generator competes with multiple discriminators  $\{D_i\}$ at different hierarchies (Figure \ref{fig:archs-review} D), each focuses on learning discriminative features in different contextual scale.

In principle, the lower-resolution side output is used to learn semantic consistent image structures (e.g. object shape, colors, and background), and the subsequent higher-resolution side outputs are used to render fine-grained details. Since our method is trained in an end-to-end fashion, the lower-resolution outputs can also fully utilize top-down knowledge from discriminators at higher resolutions. As a result, we can observe consistent image structures, color and styles in outputs of both low and high resolutions without using addition regularization terms.

%Since our method is trained in an end-to-end manner, the lower-resolution side outputs can fully enjoy the top-down knowledge from higher resolutions, such that we can observe consistent image structures and styles appear in both low and high resolutions.
%The experiment demonstrates this advantage compared with StackGAN. 
% % According to our empirical observation, our design is able to reduce the model instability, especially at the early stage of the training.
%% %We present the following multi-purpose adversarial loss to support more effective generator training.

\subsection{Multi-purpose adversarial losses}
Our generator produces resolution-growing side outputs composing an image pyramid. 
We leverage this hierarchy property and allow adversarial losses to capture hierarchical image statistics, with a goal to guarantee both semantic consistency and image fidelity.

In order to guarantee semantic consistency, we adopt the matching-aware pair loss proposed by \cite{reed2016generative}. The discriminator is designed to take image-text pairs as inputs and be trained to identify two types of errors: a real image with mismatched text and a fake image with conditioned text.

The pair loss is designed to guarantee the global semantic consistency. However, there is no explicit loss to guide the discriminator to differentiate real images from fake images. And combining both tasks (generating realistic images and matching image styles with text) into one network output complicates the already challenging learning tasks. Moreover, as the image resolution goes higher, it might be challenging for a global pair-loss discriminator to capture the local fine-grained details (results are validated in experiments).
%As the image resolution goes higher, discriminators might loss focus for local fine-grained image details.
%However, this loss has two major weaknesses. First, there is no explicit loss that force the discriminator to differentiate real images from fake images. 
%Combining both tasks (generating realistic images and matching image with text) into one output further complicates the learning tasks, which is already challenging. 
%Second, as the image resolution goes higher, it might be challenging for a global pair-loss discriminator to capture local fine-grained details.  
In addition, as pointed in \cite{shrivastava2016learning}, a single global discriminator may over-emphasize certain biased local features and lead to artifacts.

To alleviate these issues and guarantee image fidelity, our solution is to add local adversarial image losses.
%as inspired by \cite{shrivastava2016learning}. 
%The insights are two folds:
%\begin{itemize}\vspace{-.1cm}
%\setlength\itemsep{-0.3em}
%\item Further forcing the generated samples to have only only global but also local similar statistics to real image patches, which is beneficial for rendering fine-grained details. 
%
%\item \textcolor{red}{Image loss without using paired text can be beneficial in the sense that it focus more on image information. The richness of discriminator supervision usually has advantages for generalization. } 
%\vspace{-.2cm}
%\end{itemize}
We expect the low-resolution discriminators to focus on global structures, while the high-resolution discriminators to focus  on local image details.
Each discriminator $D_i$ consists of two branches (see Section 3.4), one computes a single scalar value for the pair loss and another branch computes  $R_i{\times}R_i$ 2D output maps $O_i$ for local image loss.
%with an extra branch of the discriminator for the pair loss (see Section 3.4).
%computes a $R{\times}R$ probability map. 
Each element in $O_i$ differentiates whether a corresponding local image patch is real or fake.
For each $D_i$, we control $R_i$ accordingly to tune the receptive field of each element of $O_i$. 
%In practice, the local image adversarial loss can be implemented as an fully CNN \cite{shrivastava2016learning,zhu2017unpaired,isola2016image}. 
Figure \ref{fig:loss} illustrates how hierarchically-nested discriminators compute the two losses on the generated images in the pyramid.

%Specifically, a larger $R_i$ value results in a smaller local range and finer focus of discriminators.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.49\textwidth]{figure/loss.pdf}
	\vspace{-.7cm}
	\caption{Given a set of images from the side output pyramid of the generator, the corresponding discriminator $D_i$ computes the matching-aware pair loss and local image loss (outputting a $R_i{\times}R_i$ probability map to classify real or fake image patches). }  \vspace{-.2cm}
	\label{fig:loss}
\end{figure}

\textbf{Full Objective } Overall, our full min-max adversarial objective can be defined as 
%%hierarchically-nested discriminators $\{D_i\}$ optimize the following objective:
\vspace{-.2cm}
\begin{equation}
\begin{split}
\mathcal{V}(& \mathcal{G},\mathcal{D}, \mathcal{Y}, \bm t, \bm z) = \sum_{i=1}^{s} \Big(  L_2\big(D_i({Y}_i)\big) +  L_2\big(D_i({Y}_i, \bm t_{Y})\big) \, + \\ 
& \overline{L_2}\big(D_i({X}_i)\big)  + \overline{L_2}\big(D_i({X}_i, \bm{t}_{X_i})\big) + \overline{L_2}\big(D_i({Y}_i,  \bm{t}_{\overline{Y}})\big) \Big),
\end{split}
\end{equation}
where $L_2(x) = \mathbb{E}\big[(x - \mathbb{I})^2\big]$ is the mean-square loss (instead of the conventional cross-entropy loss) and $\overline{L_2}(x) =\mathbb{E}\big[x^2\big]$. 
This objective is directly minimized by $\mathcal{D}$. For generator $\mathcal{G}$ in practice, we minimize 
$\sum_{i=1}^{s} (L_2(D_i(\mathcal{G}(\bm t, \bm z)_i)) +  L_2\big(D_i(\mathcal{G}(\bm t, \bm z)_i, \bm{t}_{X_i})))$ instead.
For the local image loss, the shape of $x, \mathbb{I} \in \mathbb{R}^{R_i{\times}R_i}$ vary accordingly (see Figure \ref{fig:loss}). $R_i=1$ refers to the (largest local) global range. $D_i(X_i)$ is the image loss branch and $D_i(X_i, \bm{t}_{X_i})$ is the pair loss branch (conditioned on $\bm{t}_{X_i}$).
$\{Y_i, \bm t_{Y}\}$ denotes a matched image-text pair and $\{Y_i, \bm{t}_{\overline{Y}}\}$ denotes a mismatched image-text pair.

In the spirit of variational auto-encoder \cite{vae} and the practice of StackGAN \cite{han2017stackgan}, instead of directly using the deterministic text embedding, we sample a stochastic vector from a Ganssian distribution $\mathcal{N}(\mu({\bm t}), \Sigma({\bm t}))$, where $\mu$ and $\Sigma$ are parameterized functions of $\bm t$. 
We add the Kullback-Leibler divergence regularization term, $D_{KL}(\mathcal{N}(\mu({\bm t}), \Sigma({\bm t}) )|| \mathcal{N}(0, \bm{I}))$, to the GAN objective to prevent over-fitting and force smooth sampling over the text embedding distribution.

%\begin{equation}
%{\small 
%\label{equ:pariG}
%\begin{split}
% \mathcal{L}(\mathcal{G}, \mathcal{D})  = & \mathbb{E}_{ t \sim p_{t}, z \sim p_{z}  , X \sim p_{\mathcal{G}(t, z)}} \Big[ \\
%&   \sum_{i=1}^{s} \big( M(D_i({X}_i)) + M(D_i({X}_i, \bm t)) \big)  + \\
%&   \sum_{i=1}^{s} \big( \hat{M}(D_i({X}_i)) + \hat{M}(D_i({X}_i, \bm t)) \big) \Big] 
%\end{split}
% }
%\end{equation}

%\begin{equation}
%\label{equ:totalG}
%\begin{split}
%\mathcal{L}_{g}(G,o,\hat{o}, t)  = &  \mathcal{L}_{pair}(G) + \frac{\alpha}{2}~(\mathcal{L}_{local}(G, o) + \mathcal{L}_{local}(G, \hat{o})) + \\
%& \beta~D_{KL}(\mathcal{N}(\mu({t}), \Sigma({t}) )|| \mathcal{N}(0, \bm{I})) 
%\end{split}
%\end{equation}

%In addition, at the very beginning of the training,  generated images are usually far away from true image distribution, the discriminator can very easily judge a pair as invalid without using much pair information. \textcolor{red}{need more explanationn.}
%We not only want the whole image is classified as true image by the discriminator, but also want to restrict the discriminator to classify all the local patches of high resolution output as true as well.

%so that they can focus more on local details, we restrict the receptive field of the image discriminator.

%\subsection{Matching Aware Pair-Loss}
%During the setting of conditional generative adversarial learning, simply generating images of high visual quality is not enough, most importantly, we want the generated images well aligned with the conditioning information (ie, text description in our model).     
%Given a pair of image and text embedding $(x, \psi{t})$ as input, in the discriminator $D_i$, we first transfer $x$  to an image code $Code_x$. $Code_x$ along with $f_{sent}(\psi_t)$ is concatenated and fed to the matching aware branch to compute a final scale value indicating the probability of the pair being real or fake, where $f_{sent}$ consists of one linear layer and one leaky ReLU transformation,  the matching-aware branch is a set of residual blocks with strid-2 convolution, more details of the architecture will be provided in the appendix. Besides that, the image code is also fed to an image-disc branch, whose structure is very similar to the matching-aware branch, to judge the probability of the input image being real or fake. We performance batch normalization on all convolutional layers. \textcolor{red}{like stackgan, I will add much more architecture details later}  
%
%The generator $G$ and discriminator $D$ are optimized alternatively using their corresponding loss function. 
%Therefore, the discriminator are designed to take image and sentence pair as input and should be able to identify two types of errors: real image with mismatch conditional text information and fake image with real text. In our case, the discriminator is explicitly trained with three pair configurations, namely: real image with right text as positive training sample, and the aforementioned two sources of error pairs as negative training samples.

%Given a mini-batch of training data consists of images $X$ and the corresponding text $t$, mismatching text $\hat{t}$. For simplicity, we omit the batch index.  To train the discriminator,  we also need to sample fakes images $o_i$ using $G$. For the $i-$th discriminator $D_i$ , we can have:

%\begin{equation}
%\label{equ:matchignloss}
%\begin{split}
%\mathcal{L}_{pair}(D_i)  = & \mathbb{E}_{(X_i, \hat{t}) \sim p_{data}}[(D_i( {X_i},  {t}) - 1) ^ 2 ] - \frac{1}{2}*\\ 
%&   \mathbb{E}_{(X_i, \hat{t}) \sim p_{data}}[(D_i({X_i}, \hat{{t}}) - 1)^2  ] - \frac{1}{2} *\\
%&   \mathbb{E}_{t \sim p_{data}}[ (D_i( {o}_i, {t}))^2 ]  
%\end{split}
%\end{equation}
%
%Similarly, the pair wise objective of generator can be summarized as:
%\begin{equation}
%\label{equ:pariG}
%\begin{split}
%\mathcal{L}_{pair}(G)  = & \sum_{i=1}^{s} ( \mathbb{E}_{t \sim p_{data}}[ (D_i( {o}_i, {t}) - 1) )^2 ] )
%\end{split}
%\end{equation}
%
%where $o_i$ is the $i-$th output of the $G$,  this type of loss is used to push the valid image-text pair away from the invalid pairs.
%
%\subsection{Local Adversarial Image Loss}
%%As is shown in Figure. \ref{overview}, the discriminator $D$ in our model mainly consists of two branches: matching-aware pair discrimination branch $D_{pair}$, and image discrimination branch $D_{img}$, both branches share the same  image encoder.
%
%The pair-loss is designed to differentiate invalid pairs from valid ones.
%It works by first translating the image to a compact size code, then the image code and the sentence vector is concatenated and fed to successive layers to predict the validity of the input image-sentence pair. 
%
%Overall, the pair loss is designed to capture the global context semantic information.  However, this loss has two problems, firstly,  there is no explicit loss that push the discriminator to differentiate real images from the fake images. Combining both tasks (generating realistic images and matching image with text) into one output further complicates the learning tasks, which is already challenging. Second,  as the image resolution goes higher, it might be challenging for the global pair wise discriminator to capture local fine-grained details.  In addition, as is also pointed in \cite{shrivastava2016learning}, a single global strong discriminator tend to over-emphasize certain image feature and may lead to artifacts. 
%
%%More specifically, ideally the image patches sampled from the higher resolution outputs 
%
%Our solutions is to add hierarchical local adversarial image loss which does not require sentence information, to all the side outputs. To make the image discriminator focus more on local details, we restrict the receptive field of the image discriminator.
%Our insight has two folds:
%\begin{itemize}
%	\item 1. Forcing the local patch of generated samples to have similar statistics to real image patch can be beneficial to rendering fine-grained details. %
%	 
%	\item 2. Image loss without using pair sentence information can be beneficial in the sense that it utilize more information as well as it directly push the generated image to the true image distribution. We hypothesis this type of loss can help the model to stabilize and boost the training. 
%\end{itemize}
%%In addition, at the very beginning of the training,  generated images are usually far away from true image distribution, the discriminator can very easily judge a pair as invalid without using much pair information. \textcolor{red}{need more explanationn.}
%%We not only want the whole image is classified as true image by the discriminator, but also want to restrict the discriminator to classify all the local patches of high resolution output as true as well.   
%Since our model generates a pyramid of outputs with increasing image sizes. 
%We expect the low resolution output capture the global structure and color information, while the higher resolution outputs focus on local image details. 
%
%To force the discriminator focus on the fine-grained details, a nature way is to fix the receptive field for all the generator. To this end, we propose a hierarchical image loss, where the output of image  discriminator is a grid of values. 
%
%Denote $K_i$ as the $i-$th local image discriminators,  the output $d_i = K_i(X_i)$ has a size increasing with input $X_i$. For instance 1$\times$1 for 64$\times$64 input, 4$\times 4$ for 128$\times$128 and 8$\times 8$ for 256$\times$256 etc.  In this way, even the resolutions goes higher, the receptive field keeps unchanged.  Specifically, the proposed local image loss can be given by:
%\begin{equation}
%\label{equ:local}
%\begin{split}
%\mathcal{L}_{local}(K_i, X_i, o_i, t )  = & \mathbb{E}_{(x, t) \sim p_{data}}[ \mathbb{E}(K_i( {X_i}) - 1) ^ 2 ] - \\
%&  \mathbb{E}_{t \sim p_{data}}[  \mathbb{E}(K_i( {o}_i))^2 ]  
%\end{split}
%\end{equation}
%
%Similarly, the pair wise objective of generator can be summarized as:
%\begin{equation}
%\label{equ:pariG}
%\begin{split}
%\mathcal{L}_{local}(G, o)  = & \sum_{i=1}^{s} ( \mathbb{E}_{t \sim p_{data}}[ \mathbb{E} (K_i( {o}_i) - 1) )^2 ] )
%\end{split}
%\end{equation}
%Further more, this separation of the image loss from the matching loss open further possibilities, such as utilizing the unlabeled image to improve both the generator and discriminator.

%
%\subsection{Total Loss}
%
%Our objective for both $D$ and $G$ thus mainly compress two types of terms: \textit{matching aware pair-loss} for matching the overall image content to the text description, and \textit{local adversarial image loss} for rendering the fine-grained details. In the spirit of variational auto-encoder and the practice of \cite{han2017stackgan}, we also add the  Kullback-Leibler divergence regularization term to the $G$ loss.  
%
%%In addition, this term might also play a role in balancing the $c_t$ and generated condition vector $c_t$ by forcing $c$ to have a similar scale to $z$.  In the following sections, we denote $c_t$ as the variational conditional text vector.
%
%\begin{equation}
%\label{equ:totalD}
%\begin{split}
%\mathcal{L}_{d}(D_i, X_i, o_i,\hat{o_i}, t, \hat{t})    = &\mathcal{L}_{pair}(D_i) + \frac{\alpha}{2}~( \\
%&\mathcal{L}_{local}(K_i, o_i) + \mathcal{L}_{local}(K_i, \hat{o_i}) )
%\end{split}
%\end{equation}
%\begin{equation}
%\label{equ:totalG}
%\begin{split}
%\mathcal{L}_{g}(G,o,\hat{o}, t)  = &  \mathcal{L}_{pair}(G) + \frac{\alpha}{2}~(\mathcal{L}_{local}(G, o) + \mathcal{L}_{local}(G, \hat{o})) + \\
%                  & \beta~D_{KL}(\mathcal{N}(\mu({t}), \Sigma({t}) )|| \mathcal{N}(0, \bm{I})) 
%\end{split}
%\end{equation}
%where $\alpha$ is coefficient used to tune the weight of the local patch loss, $\beta$ is used to tune the weights of the KL-Divergence. 
%We omit the additional arguments to both $\mathcal{L}_{pair}(G) $ and $\mathcal{L}_{pair}(D_i) $ for simplicity.
%The loss function of the matching aware discriminator are summarized as:
%\begin{equation}
%\label{matchignloss}
%\begin{split}
%\mathcal{L}_D = 	  &\mathbb{E}_{(x, t) \sim p_{data}}[\log( D( x  \psi_t) ) ] -  0.5*(\\
%&\mathbb{E}_{(x, \hat{t}) \sim p_{data}}[\log(D(x, \psi_{\hat{t}} )  ] +\\
%&   \mathbb{E}_{t \sim p_{data}}[\log(D(\hat{x_1}, \psi_t) ) ] 
%\end{split}
%\end{equation}
%
%\begin{equation}
%\label{matchignloss}
%\begin{split}
%\mathcal{L}_D = 	  &\mathbb{E}_{(x, t) \sim p_{data}}[\log( D( x  \psi_t) ) ] -  0.5*(\\
%&\mathbb{E}_{(x, \hat{t}) \sim p_{data}}[\log(D(x, \psi_{\hat{t}} )  ] +\\
%&   \mathbb{E}_{t \sim p_{data}}[\log(D(\hat{x_1}, \psi_t) ) ] 
%\end{split}
%\end{equation}

% \begin{algorithm}
% 	\caption{Multi scale GAN training algorithm with step size $\alpha$, number of iteration $S$, we omit the batch index and summarize the algorithm using SGD for simplicity.}
% 	\label{CHalgorithm}
% 	\begin{algorithmic}[1]
% 		\For{$i \gets 1 \textrm{ to } S$ }
% 		\For{$k \gets 1 \textrm{ to } ncritic$}
% 		\State Sample two noise batch $z_1, z_2 \sim p(z)$.
% 		\State Sample one batch of matching and mismatching image text pair $\{X, t\}$,  $\{X, \hat{t}\}$.
% 		\State Generate fake image samples for both $t$ and $\hat{t}$
% 		\begin{equation}
% 		\begin{split}
%	 		& {o} = G(t, z_1) \\
%	 		& {\hat{o}} = G(\hat{t}, z_2) 
% 		\end{split}
% 		\end{equation}
%	 		\For{each discriminator $D_i$ }
%		 		\State Update $D_i$ by descending the gradient of loss \ref{equ:totalD}:
%		 		\begin{equation}
%		 		   	\theta_{D_i} = \theta_{D_i}  - \nabla_{\theta_{D_i}} \mathcal{L}_d(D_i, X_i,o_i,\hat{o_i}, t,\hat{t})
%		 		\end{equation}
%	 		\EndFor
% 		\EndFor
% 		\State Update the generator by descending it's stochastic gradient of loss \ref{equ:totalG}:
% 		\begin{equation}
% 		\theta_{G} = \theta_{G}  - \nabla_{\theta_{G}} \mathcal{L}_g(G, o, \hat{o}, t)
% 		\end{equation}
% 		
% 		
% 		\EndFor
% 		
% 	\end{algorithmic}
% \end{algorithm}

\subsection{Architecture Design}
%We introduce framework details as well as motivations.

\textbf{Generator} The generator is simply composed of three kinds of modules, termed as $K$-repeat res-blocks, stretching layers, and linear compression layers.
A single res-block in the $K$-repeat res-block is a modified\footnote{We remove ReLU after the skip-addition of each residual block, with an intention to reduce sparse gradients.} residual block \cite{he2016identity}, which contains two convolutional (conv) layers (with batch normalization (BN) \cite{ioffe2015batch} and ReLU). The stretching layer serves to change the feature map size and dimension. It simply contains a scale-$2$ nearest up-sampling layer followed by a conv layer with BN+ReLU. The linear compression layer is a single conv layer followed by a Tanh to directly compress feature maps to the RGB space (as side outputs). We prevent any non-linear function in the compression layer that could impede the gradient signals. 
%, whose compact design enforces feature maps in conv blocks can linearly represent RGB images at arbitrary scales.
Starting from a $1024{\times}4{\times}4$ embedding, which is computed by CA and a trained embedding matrix, the generator simply uses $M$ $K$-repeat res-blocks connected by $M{-}1$ in-between stretching layers until the feature maps reach to the targeting resolution. 
For example, for $256{\times}256$ resolution and $K{=}1$, there are $M{=}6$ $1$-repeat res-blocks and $5$ stretching layers. 
%Such simple and symmetric design makes our generator helpful for gradient flows. We will verify in experiments. 
At pre-defined side-output positions at scales $\{1,...,s\}$, we apply the compression layer to generate side output images, for the input of the discriminator.

\textbf{Discriminator} The discriminator simply contains consecutive stride-2 conv layers with BN+LeakyReLU. There are two branches are added to the upper layer of the discriminator. One branch is a direct fully convolutional layers to produce a $R{\times}R$ probability map (see Figure \ref{fig:loss}) and classify each location as real or fake. 
Another branch first concatenates a $512{\times}4{\times}4$ feature map and a $128{\times}4{\times}4$ text embedding (replicated by a reduced $128$-d text embedding). Then we use an $1{\times}1$ conv to fuse text and image information and a $4{\times}4$ conv layer to classify the image-text pair to real or fake.
%%it is critical to guarantee the semantic consistency in the final results.

The optimization is similar to the standard alternative training strategy in GANs. Please refer to the supplementary material for more training and network details.

%for the consideration of memory saving and faster convergence.

%\begin{equation}
%\label{genloss}
%\begin{split}
%\mathcal{L}_G = & \mathbb{E}_{z\sim p_{z}, t \sim p_{data}}[-\log( D_{sent}( G(z, c_t),  \psi_t) ) ] + \\
%& \mathbb{E}_{z\sim p_{z}, t \sim p_{data}}[-\log( D_{img}( G(z, c_t) ) ) ] + \\
%& D_{KL}(\mathcal{N}(\mu(\psi_t), \Sigma(\psi_t) )|| \mathcal{N}(0,` \bm{I})) 		   
%\end{split}
%\end{equation}
%where $c_t$ is the variational conditional vector generated from $\mathcal{N}(\mu(\psi_t), \Sigma(\psi_t) )$. To make the network differentiable, in practice, we adopted the re-parametric trick\cite{vae}, where $c_t$ is actually computed as $\mu(\psi_t)+\Sigma(\psi_t)\odot \epsilon$, where $\epsilon\in \mathcal{R}^{T}$ is draw from a normal Gaussian distribution.

%\begin{algorithm}
%	\caption{Multi scale GAN training algorithm with step size $\alpha$, number of iteration $S$, we summarize the algorithm using SGD for simplicity.   
%		\label{alg:gan-train}}
%	\begin{algorithmic}[1]
%		\Input{sample a minibatch of images, $x$, matching text description $t$, mis-matching $\hat{t}$,
%			}
%		\For{$i \gets 1 \textrm{ to } S$}
%			$z_1, z_2 \sim \mathcal{N}(0,\bm{1})^Z$
%			\State $\hat{x_1}$=$G(z_1, \psi_t)$ \Comment{Forward through $G$ using matching text embedding}
%			\STATE {$\hat{x_2}$}{$G(z_2, \psi_\hat{t})$} \Comment{Forward through $G$ using mis-matching text embedding}
%			
%			\STATE {$d_r$}{$D_{pair}(x, \psi_t)$} \Comment{(real image, matching text)}
%			\STATE {$d_m$}{$D_{pair}(x, \psi_\hat{t})$} \Comment{(real image, mis-matching text)}
%			\STATE {$d_f$}{$D_{pair}(\hat{x_1}, \psi_t)$} \Comment{(fake image, matching text)}
%			\STATE {$s_r$}{$D_{img}(x)$} \Comment{Forward through $D_{img}$ with $x$}
%			\STATE {$s_1$}{$D_{img}(\hat{x_1})$} \Comment{Forward through $D_{img}$ with $\hat{x_1}$}
%			\STATE {$s_2$}{$D_{img}(\hat{x_1})$} \Comment{Forward through $D_{img}$ with $\hat{x_2}$}
%			\STATE {$\mathcal{L}_{D}$}{$(d_r-1)^2 + (s_r-1)^2 + (d_m + d_f + s_1 + s2)/2 $} 
%			\STATE {$\mathcal{L}_{G}$}{$((d_f -1)^2 + (s_1 -1)^2 + (s_2 -1)^2 $} 
%			\STATE {$\theta_G$}{$\frac{\delta \mathcal{L}_G}{\delta \theta_G}$}  \Comment{Update generator}
%			\STATE {$\theta_D$}{$\frac{\delta \mathcal{L}_D}{\delta \theta_D}$}  \Comment{Update generator}
%		\EndFor	
%	\end{algorithmic}
%\end{algorithm}

\section{Experiments}
%%This section evaluates the proposed method both qualitatively and quantitatively on three public datasets. 
We denote our method as \textbf{HDGAN}, referring to High-Definition results and the idea of Hierarchically-nested Discriminators.

\textbf{Dataset} 
We evaluate our model on three widely used datasets. The CUB dataset \cite{welinder2010caltech} contains 11,788 bird images belonging to 200 categories. 
%, which crops the bird region and ensure the foreground occupies greater than 0.75.  
The Oxford-102 dataset \cite{Nilsback08} contains 8,189 flow images in 102 categories. 
%We follow \cite{han2017stackgan, reed2016generative} to split both of these datasets into class-disjoint training and testing sets. 
Each image in both datasets is annotated with 10 descriptions provided by \cite{reed2016generative}. We pre-process and split the images of CUB and Oxford-102 following the same pipeline in \cite{reed2016generative,han2017stackgan}. The COCO dataset \cite{lin2014microsoft} contains 82,783 training images and 40,504 validation images. Each image has 5 text annotations. 
We use the pre-trained charRNN text encoder provided by \cite{reed2016generative} to encode each sentence into a 1024-d text embedding vector.
%During the training stage, we randomly pick-up a cropped view of the original images, and pick 4 image sentence embedding and use the mean as conditioning vector.

\textbf{Evaluation metric}
We use three kinds of quantitative metrics to evaluate our method.
%\textcolor{red}{\textbf{Inception score.}}  To evaluate the quality of generated images, we resort to the recently proposed qualitative evaluation measurement "Inception score"\cite{improvedGAN},
% \begin{equation}
%	IS = \exp(\mathbb{E}_xD_{\bm{KL}}(p(y|\bm{x})||p(y)))
% \end{equation}
% where $\bm{x}$ denotes generated sample images, $y$ the class label, $p(y|\bm{x})$ represents the conditional label distribution calculated by the Inception model\cite{Inception}. 
1) Inception score \cite{improvedGAN} is a measurement of both objectiveness and diversity of generated images. Evaluating this score needs a pre-trained Inception model~\cite{inception} on ImageNet. For CUB and Oxford-102, we use the fine-tuned Inception models on the training sets of the two datasets, respectively, provided by StackGAN. 
2) Multi-scale structural similarity (MS-SSIM) metric \cite{improvedGAN} is used for further validation. It tests pair-wise similarity of generated images and can identity mode collapses reliably \cite{odena2016conditional}. Lower score indicates higher diversity of generated images (i.e. less model collapses).

3) \textbf{Visual-semantic similarity} The aforementioned metrics are widely used for general GANs. However, they can not measure the alignment between generated images and the conditioned text, i.e., semantic consistency. \cite{han2017stackgan} resorts to human evaluation, but this procedure is expensive and difficult to conduct.
To tackle this issue, we introduce a new qualitative measurement, namely visual-semantic similarity (VS similarity) inspired by \cite{vsemb}.
Denote $\bm{v}$ as the image feature vector extracted by a pre-trained Inception model $f_{cnn}$.
We define a scoring function $c(\bm{x}, \bm{y})=\frac{\bm{x}\cdot \bm{y}}{||\bm{x}||_2\cdot ||\bm{y}||_2}$. 
Then, we learn two mapping functions $f_v$ and $f_t$, which map both image and sentence embeddings into a common space in $\mathbb{R}^{512}$, by minimizing the following bi-directional ranking loss:
%\mathcal{L}_{vs} &=
\begin{equation}\vspace{-.2cm}
\begin{split}
  \sum_{ \bm{v} }\sum_{ \bm{t}_{\overline{v}} }~ &\max( 0, \delta  - c(f_v(\bm{v}), f_t( \bm{t}_{{v}}) )+ c(f_v(\bm{v}), f_t( \bm{t}_{\overline{v}}) )     ) +\\
   \sum_{ \bm{t} }\sum_{ \bm{v}_{\overline{t}} }~ &\max( 0,  \delta  - c(f_t(\bm{t}), f_v( \bm{v}_{{v}} )) +  c(f_t(\bm{t}), f_v( \bm{v}_{\overline{t}} ))  ),
\end{split}
\vspace{-.2cm}
\end{equation}
where $\delta$ is the margin, which is set as 0.2. $\{\bm{v}, \bm{t}\}$ is a ground truth image-text pair, and $\{\bm v, \bm t_{\overline{v}} \}$ and $\{ \bm v_{\overline{t}}, \bm{t} \}$ denote mismatched image-text pairs. In the testing stage, given an text embedding $\bm{t}$, and the generated image $\bm{x}$, the VS score can be calculated as $c(f_{cnn}(\bm{x}), \bm{t})$. Higher score indicates better semantic consistency.

\begin{table}[t] % retrieval
	\begin{center}
		\small 
		\begin{tabularx}{.466\textwidth}{c|ccc}

			\specialrule{1.5pt}{0pt}{0pt}  
			\multirow{2}{*}{Method}	& \multicolumn{3}{c}{Dataset}	\\ \cline{2-4}
							 		&	 CUB		&	Oxford-102  & COCO		     \\ \hline
			GAN-INT-CLS 	&	$2.88{\pm}.04$		& 	$2.66{\pm}.03$		& $7.88{\pm}.07$	 \\
			GAWWN 	  &		$3.60{\pm}.07$		&     -      &          - \\ 
			StackGAN     &		$3.70{\pm}.04$	&	 $3.20{\pm}.01$			&  $8.45{\pm}.03^{\star}$		\\ 
		%%	StackGAN$^\star$     &		$3.89{\pm}.05$	&	 $3.16{\pm}.01$			&  $8.45{\pm}.03$		\\ \hline
			StackGAN++     &		$3.84{\pm}.06$	&	 -			&  -	\\  
			TAC-GAN	 &	-		&		$3.45{\pm}.05$		& -	\\	\hline
%			HDGAN 		&	$\bm{4.15{\pm}.05}$	&	$ \bm{3.45{\pm}.07}$	&  $ \bm{11.29{\pm}.18}$  \\ \hline
			HDGAN 		&	$\bm{4.15{\pm}.05}$	&	$ \bm{3.45{\pm}.07}$	&  $ \bm{11.86{\pm}.18}$  \\ \hline
		\end{tabularx} 
	\end{center}
	\vspace{-.4cm}
	\begin{tablenotes}
		\small
		\item $^\star$Recently, it updated to ${10.62{\pm}.19}$ in its source code.
	\end{tablenotes} \vspace{-.1cm}
	\caption{The Inception-score comparison on the three datasets. HDGAN outperforms others significantly.} \label{table:score}
\end{table}

\begin{figure*}[t]
	\centering
	\includegraphics[width=0.99\textwidth]{figure/cub.pdf}
	\vspace{-.2cm}
	\caption{Generated images on CUB compared with StackGAN. Each sample shows the input text and generated $64^2$ (left) and $256^2$ (right) images. Our results have significantly higher quality and preserve more semantic details, for example, ``the brown and white feathers and read head'' in the last column is much better reflected in our images. Moreover, we observed our birds exhibit nicer poses (e.g. the frontal/back views in the second/forth columns). Zoom-in for better observation.}
	\vspace{-0.3cm}
	\label{fig:vis-cub}
\end{figure*}
\begin{figure*}[t]
	\centering
	\includegraphics[width=0.99\textwidth]{figure/flowers.pdf}
	\vspace{-.2cm}
	\caption{Generated images on Oxford-102 compared with StackGAN. Our generated images perform more natural satisfiability and higher contrast and can generate complex flower structures (e.g. spiked petals in the third example).} \label{fig:vis-oxford}
	\vspace{-.5cm}
\end{figure*}

\begin{table}[t] % retrieval
	\small
	\begin{center}
		\begin{tabularx}{.475\textwidth}{c|ccc}
			\specialrule{1.5pt}{0pt}{0pt}  
			\multirow{2}{*}{Method}	& \multicolumn{3}{c}{Dataset}	\\ \cline{2-4}
			&	 CUB		&	Oxford-102  & COCO		     \\ \hline
			Ground Truth	&	${.302{\pm}.151}$	&	$ {.336{\pm}.138}$			& $.426{\pm}.157$  \\ \hline
			StackGAN     &	$.228{\pm}.162$	&	 $.278{\pm}.134$			&  $-$		\\ 
			HDGAN 		&	$\bm{.246{\pm}.157}$	&	$ \bm{.296{\pm}.131}$ & $\bm{.199{\pm}.183}$  \\ \hline
		\end{tabularx} 
	\end{center}
	\vspace{-.4cm}
	\caption{The VS similarity evaluation on the three datasets. The higher score represents higher semantic consistency between the generated images and conditioned text. The groundtruth score is shown in the first row.} \label{table:vss} \vspace{-.3cm}
\end{table}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.48\textwidth]{figure/coco2.pdf}
	\vspace{-.6cm}
	\caption{Samples on the COCO validation set, which contain descriptions across different scenes. } \label{fig:coco}
	\vspace{-.7cm}
\end{figure}

\subsection{Comparative Results}
To validate our proposed HDGAN, we compare our results with GAN-INT-CLS \cite{reed2016generative}, GAWWN \cite{reed2016learning}, TAC-GAN \cite{dash2017tac},  Progressive GAN \cite{Karras2017progressive}, StackGAN \cite{han2017stackgan} and also its improved version StackGAN++ \cite{han2017stackganv2}\footnote{StackGAN++ and Prog.GAN are two very recently released preprints we noticed. We acknowledge them as they also target at generating high-resolution images. }. We especially compare with StackGAN in details (results are obtained from its provided models).

Table \ref{table:score} compares the Inception score. We follow StackGAN' experiment setting to sample ${\sim}30,000$ $256^2$ images for evaluation.
HDGAN achieves significantly improvement compared against other methods. For example, it significantly improves StackGAN by $.45$ and StackGAN++ by $.31$ on CUB.
HDGAN achieves competitive results with TAC-GAN on Oxford-102. TAC-GAN uses image labels to increase the discriminability of generators, while we do not use any extra knowledge. Figure \ref{fig:vis-cub} and Figure \ref{fig:vis-oxford} compare the qualitative results with StackGAN on CUB and Oxford-102, respectively, by demonstrating more, semantic details, natural color, and complex object structures. 
Moreover, we qualitatively compare the diversity of samples conditioned on the same text (with random input noises) in Figure \ref{fig:multiple-test} left. HDGAN can generate obviously more compelling samples. See figure captions for explanations. The demonstrated results strongly prove the superiority of HDGAN.

Different from CUB and Oxford-102, COCO is much more challenging and contains largely diverse natural scenes. 
Our method obtains significantly improvement on COCO (Table \ref{table:score}). 
Figure \ref{fig:coco} also shows some generated samples in several different scenes.
Please refer to the supplementary material for more results.

Table \ref{table:vss} compares the proposed visual semantic similarity (VS) results on three datasets. The score of the groundtruth image-text pair are also shown for reference.  HDGAN achieves consistently better performance on both CUB and Oxford-102. These results demonstrate that HDGAN can better capture the visual semantic information in generated images.

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=0.499\textwidth]{figure/multisample_cub.pdf}
%	\caption{Results on Oxford} \label{fig:multires-cub}
%\end{figure}

\begin{figure*}[t]
	%	\begin{center}
	\centering
	\begin{subfigure}[t]{0.69\textwidth}
		%	\centering
		\includegraphics[width=0.99\textwidth]{figure/multisample_cub.pdf}
	\end{subfigure} 
	\begin{subfigure}[t]{0.3\textwidth}
		%	\centering
		\includegraphics[width=0.99\textwidth]{figure/cub_multires.pdf}
	\end{subfigure}
	\vspace{-.2cm}
	\caption{Left: Diverse samples are shown given a single input text. The proposed HDGAN (top) show obviously more fine-grained details. Right: Side outputs of HDGAN with increasing resolutions. Different resolutions are semantically consistent and semantic details appear as the resolution increases.  \label{fig:multiple-test}} 	\vspace{-.3cm}
	%\end{center}
\end{figure*}

\begingroup
\setlength{\intextsep}{-4pt}%
\setlength{\columnsep}{0pt}%
\begin{wrapfigure}{r}{0.24\textwidth}
	\centering
	\includegraphics[width=0.245\textwidth]{figure/multiscale_inception_2.pdf}
	\vspace{-12pt}
\end{wrapfigure}
%Table \ref{table:multiscales}
The right figure compares the multi-resolution Inception score on CUB. Our results are from the side outputs of a single model. As can be observed, our $64^2$ results outperform the $128^2$ results of StackGAN and our $128^2$ results also outperform $256^2$ results of StackGAN substantially. It  demonstrates that our HDGAN better preserves semantically consistent information in all resolutions (as stated in Section 3.2). Figure \ref{fig:multiple-test} right validate this property qualitatively. On the other hand, we observe that, in StackGAN, the low-resolution images and high-resolution images sometimes are visually inconsistent with respect to colors and structures (see examples in Figure \ref{fig:vis-cub} and Figure \ref{fig:vis-oxford}).

Table \ref{fig:msssmi} compares the MS-SSMI score with StackGAN and Prog.GAN for bird image generation. StackGAN and our HDGAN use text as input so the generated images are separable in class. We randomly sample ${\sim}20,000$ image pairs (400 per class) and compare the class-wise score in the left figure. HDGAN outperforms StackGAN in majority of classes and also has a lower standard deviation ($.023$ vs. $.032$).
Note that Prog.GAN uses noise input rather than text. We can compare with it for the general measure of the image diversity. Following the procedure of Prog.GAN, we randomly sample ${\sim}10,000$ image pairs from all generated samples\footnote{We use $256^2$ bird images provided by Prog.GAN at \url{https://github.com/tkarras/progressive_growing_of_gans}. Note that Prog.GAN is trained on the LSUN \cite{yu2015lsun} bird set, which contains ${\sim}2$ million bird images.} and show the results in Table \ref{fig:msssmi} right. HDGAN outperforms both methods. 
\begin{table}[t] % retrieval
	\begin{minipage}[b]{0.50\linewidth}
		%\centering
		\includegraphics[width=0.99\textwidth,height=0.7\textwidth]{figure/ms_ssmi.pdf}
		\vspace{-1.8cm}
	\end{minipage} %\hfill
	\begin{minipage}[b]{0.49\linewidth}
		\begin{tabularx}{.9\textwidth}{c|c}
			\specialrule{1.5pt}{0pt}{0pt}  
			Method   &  MS-SSMI \\ \hline
			StackGAN &   0.234   \\ 
			Prog.GAN &   0.225    \\ \hline
			HDGAN    &   \textbf{0.215}    \\ \hline
		\end{tabularx}
	\end{minipage}
	\vspace{0.2cm}
	\caption{Left: Class-wise MS-SSMI evaluation. Lower score indicates higher intraclass dissimilarity. The points below the equality line represent classes our HDGAN wins. The inter-class std is shown in axis text. Right: Overall (not class-wised)  MS-SSMI evaluation.} \label{fig:msssmi}
	\vspace{-0.4cm}
\end{table}

\subsection{Style Transfer Using Sentence Interpolation}
Ideally, a well trained model should learn a smooth linear latent manifold of the images. To demonstrate the generalization capability of our proposed HDGAN, we generate images using the linearly interpolated embeddings between two source sentences. 
As shown in Figure~\ref{fig:interp}, the generated images show smooth style transformation and well reflect the semantic details in sentences. 
%For example, in the first row, the generated birds gradually turn the color from yellow to gray. In the second row, 
For example, in the second row, complicated sentences with detailed appearance descriptions (e.g.,pointy peaks and black wing) are used, our model could still successfully capture these subtle features and tune the birds' appearance smoothly.

%In the first row, the generated birds gradually turn the color from yellow to brown. In the second row, more complicated sentences with detailed appearance descriptions (e.g., blue peaks and yellow wing) are used, we can see that our model is still able to successfully capture these subtle features and tune the birds' appearance gradually.

%\subsubsection{What role does noise play}
%In our model, there are two kinds of randomness injected to the input of the Generator, ie, epsilon $\epsilon$ and Gaussian noise vector $z$. 
%In this experiment, we fix the sentence embedding and one noise, and change the other random variable. For each random noise, we generate 4 different values. So In total 16 images are generated using all noise combinations, the resulting images are shown in Figure~\ref{}, $\epsilon$ and $z$ corresponds to row and column dimension, respectively. As is shown in the image, we can see that, $\epsilon$ can change the pose of the image, while $z$ is mainly responsible for the background variations. This clearly demonstrate that, our model learns to disentangle the semantic information from the pose and background information.

\subsection{Ablation Study and Discussion}
%%\subsubsection{Hierarchically-nested adversarial training}
\textbf{Hierarchically-nested adversarial training} Our hierarchically-nested discriminators play a role of regularizing the layer representations (at scale $\{64, 128, 256\}$). 
%Our full method applies it on feature maps at $\{64, 128, 256\}$. 
In Table \ref{table:multiscales}, we compare the performance of HDGAN by removing part of discriminators on both CUB and COCO datasets. 
As can be seen, increasing the usage of discriminators at different scales have positive effects. And using discriminators at $64^2$ is critical (by comparing 64-256 and 128-256 cases). For now, it is uncertain if adding more discriminators and even on lower resolutions would be helpful. Further validation will be conducted.
StackGAN emphasizes the importance of using text embeddings not only at the input but also with intermediate features of the generator, by showing a large drop from $3.7$ to $3.45$ without doing so. While our method only uses text embeddings at the input. Our results strongly demonstrate the effectiveness of our hierarchically-nested adversarial training to maintain such semantic information and a high Inception score.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.48\textwidth]{figure/interp.pdf}
	\vspace{-.6cm}
	\caption{Text embedding interpolation from the source to target sentence results in smooth image style changes to match the targeting sentence. } \label{fig:interp}
	\vspace{-.3cm}
\end{figure}

%\begin{table}[t] % retrieval
%	\small
%	\begin{minipage}[b]{0.49\linewidth}
%		\begin{tabularx}{1\textwidth}{c@{\hskip1.5pt}|c@{\hskip3pt}|c@{\hskip0.5pt}}
%			\specialrule{1.5pt}{0pt}{0pt}  
%			 			&   Inc. score 	 & VS   	\\  \hline
%			\thead{w/o \\ local}        &   $3.12{\pm}.02$      & $.263{\pm}.130$           \\  \hline
%			\thead{w/ \\ local}        &   $3.45{\pm}.07$ 	 &   $.296{\pm}.130$		    \\ \hline
%		\end{tabularx}
%		\vspace{-0.995cm}
%	\end{minipage} %\hfill
%	\begin{minipage}[b]{0.49\linewidth}
%		\begin{tabularx}{.98\textwidth}{ccc|c}
%			\specialrule{1.5pt}{0pt}{0pt}  
%			\multicolumn{3}{c|}{Components}	&  \multirow{2}{*}{\thead{Inception \\ score}}	\\ \cline{1-3}
%			64	& 128	& 256 			& 		\\ \hline
%			&  		&	\checkmark	&	${3.52{\pm}.04}$	\\ 
%			&  	\checkmark	&	\checkmark	&	${3.99{\pm}.04}$	\\
%			\checkmark	&  			&	\checkmark	&  ${4.14{\pm}.03}$		\\
%			\checkmark	&  \checkmark		&	\checkmark	&	${4.15{\pm}.05}$ \\ \hline
%			
%		\end{tabularx}
%	\end{minipage}
%	\vspace{-0.2cm}
%	\caption{Ablation study on the local adversarial loss (left table) and hierarchically-nested adversarial discriminators (right table). \checkmark indicates whether a discriminator at a certain scale is used. See text for detailed explanation.} \label{tab:ablation} \vspace{-0.3cm}
%\end{table}
\begin{table}[t] % retrieval
	\begin{center}
		\begin{tabularx}{.36\textwidth}{ccc|cc}
			\specialrule{1.5pt}{0pt}{0pt}  
			\multicolumn{3}{c|}{Components}	&  \multicolumn{2}{c}{Inception score}	\\ \hline
			64	& 128	& 256 			& 	CUB		&	 COCO \\ \hline
			&  		&	\checkmark	&	${3.52{\pm}.04}$& -	\\ 
			&  	\checkmark	&	\checkmark	&	${3.99{\pm}.04}$&-	\\
			\checkmark	&  			&	\checkmark	&  ${4.14{\pm}.03}$	& ${11.29{\pm}.18}$	\\
			\checkmark	&  \checkmark		&	\checkmark	&	${4.15{\pm}.05}$& ${11.86{\pm}.18}$ \\ \hline

		\end{tabularx}
	\end{center} \vspace{-.5cm}
	\caption{Ablation study of hierarchically-nested adversarial discriminators on CUB and COO. $\checkmark $ indicates whether a discriminator at a certain scale is used. See text for detailed explanation.} \label{table:multiscales} \vspace{-.3cm}
\end{table}

%

%\begin{table}[t] % retrieval
%	\small
%	\begin{minipage}[b]{0.49\linewidth}
%		\begin{tabularx}{1\textwidth}{c@{\hskip1.5pt}|c@{\hskip1pt}|c@{\hskip0.5pt}}
%			\specialrule{1.5pt}{0pt}{0pt}  
%			&   Inc. score 	 & VS   	\\  \hline
%			\thead{w/o \\ local}        &   $3.12{\pm}.02$      & $.263{\pm}.130$           \\  \hline
%			\thead{w/ \\ local}        &   $3.45{\pm}.07$ 	 &   $.296{\pm}.130$		    \\ \hline
%		\end{tabularx}
%		\vspace{-1.995cm}
%	\end{minipage} %\hfill
%	\begin{minipage}[b]{0.5\linewidth}
%			\centering
%			\includegraphics[width=0.245\textwidth]{figure/imgloss.pdf}
%	\end{minipage}
%	\vspace{-0.2cm}
%	\caption{Ablation study on the local adversarial loss (left table) and hierarchically-nested adversarial learning (right table). See text for detailed explanation.} \label{tab:ablation}
%\end{table}

%\subsection{On the Local Image Loss}

\textbf{The local image loss} We analyze the effectiveness of the proposed local adversarial image loss. 
Table \ref{table:multiscales} compares the case without using it (denoted as `w/o local image loss').
The local image loss helps improve the visual-semantic matching evidenced by a higher VS score. We hypothesis that it is because adding the separate local image loss can offer the pair loss more ``focus'' on learning the semantic consistency. 
Furthermore, the local image loss helps generate fine-grained image details. As demonstrated in Figure \ref{fig:vis-imgloss}, although both models can successfully capture the semantic details in text, the `w/ local' model generate complex object structures described in the conditioned text more precisely.

%\begin{table}[t] % retrieval
%	\begin{center}
%		\begin{tabularx}{.268\textwidth}{ccc|c}
%			\specialrule{1.5pt}{0pt}{0pt}  
%			\multicolumn{3}{c|}{Components}	&  \multirow{2}{*}{Score}	\\ \cline{1-3}
%			64	& 128	& 256 			& 		\\ \hline
%			&  		&	\checkmark	&	${3.52{\pm}.04}$	\\ 
%			&  	\checkmark	&	\checkmark	&	${3.99{\pm}.04}$	\\
%			\checkmark	&  			&	\checkmark	&  ${4.14{\pm}.03}$		\\
%			\checkmark	&  \checkmark		&	\checkmark	&	${4.15{\pm}.05}$ \\
%			
%		\end{tabularx}
%	\end{center} \vspace{-.4cm}
%	\caption{Ablation study for hierarchically-nested adversarial supervision.} \label{table:deep-nest}
%\end{table}

%\begin{table}[t] % retrieval
%	\label{tab:local_loss}
%	\begin{center}
%		\begin{tabularx}{.43\textwidth}{c|ccc}
%			\specialrule{1.1pt}{0pt}{0pt}  
%			Metric 			 		&   wo/Img 	 & w/global     &   w/local	\\  \hline
%			MS-SSMI                 &   $-$      &   $-$           &	$.215$ \\  \hline
%			VS similarity           &   $-$      &   $-$           &	$.246{\pm}.157$ \\ \hline
%			Inception score         &   $-$ 	 &   $-$		   &    $4.15{\pm}.05$ \\
%		\end{tabularx}
%	\end{center} \vspace{-.4cm}
%	\caption{Results using three quantitative evaluation metrics on birds dataset.} \label{table:imgloss}
%\end{table}

%%\subsection{Design principles}
%Note that for text-to-image synthesis, StackGAN is the only successful method to generate $256{\times}256$ images.
\textbf{Design principles} StackGAN claims the failure of directly training a vanilla $256{\times}256$ GAN to generate meaningful images. 
We test this extreme case using our method by removing all nested discriminators. Our method still generates fairly meaningful results (the first row of Table \ref{table:multiscales}), which demonstrate the effectiveness of our proposed framework (see Section 3.4).

Initially, we tried to share the top layers of the hierarchical-nested discriminators of HDGAN inspired by \cite{liu2017unsupervised}. The intuition is that all discriminators have a common goal to differentiate real and fake despite difficult scales and such sharing would reduce their inter-variances. 
However, we did not observe benefits from this mechanism and our independent discriminators can be trained fairly stably.

%\textbf{Advantages}
%The advantages of our hierarchical-nested adversary are clearer from different views, compared with the popular stacking GAN strategy \cite{}. 
%First, it is memory efficient. 
%When viewing it as ensembled multi-scale GAN (Figure \ref{}), each GAN fully uses the parameters of its lower resolution GAN. Usually, a deeper generator is necessary to compact a discriminator. Since the depth of discriminator increases as image resolution increases, if stacking separate GANs based on low resolution GAN outputs, much more layers are needed in generators as wells. While in our method, for instance, from a $64^2$ GAN to a $128^2$ GAN, only a minimum two conv layers ($1$-repeat res-blocks) are needed. 
%
%Second, it fully enjoys top-down knowledge. 
%The low resolution can not use high resolution information in the stacking GAN strategy. While in our method, it is easy to see the low resolution GAN enjoys knowledge from higher resolution. As an evidence, our $64^2$ synthetic images have higher Inception score than $64^2$ and even $128^2$ images of StackGAN (see Section 5).

\section{Conclusion}
In this paper, we present a novel and effective method to tackle the problem of generating images conditioned on text descriptions. We explore a new dimension of playing adversarial games along the depth of the generator using the hierarchical-nested adversarial objectives. A multi-purpose adversarial loss is adopted to help render fine-grained image details.
We also introduce a new evaluation metric to evaluate the semantic consistency between generated images and conditioned text.
Extensive experiment results demonstrate that our method, namely HDGAN, can generate high-resolution photographic images and performs significantly better than existing state of the arts on three public datasets.
\begin{table}[t] % retrieval
	\small
	\centering
	\begin{tabularx}{0.37\textwidth}{c|c|c}
		\specialrule{1.5pt}{0pt}{0pt}  
		&   Inc. score 	 & VS   	\\  \hline
		w/o  local image loss      &   $3.12{\pm}.02$      & $.263{\pm}.130$           \\  \hline
		w/  local  image loss      &   $3.45{\pm}.07$ 	 &   $.296{\pm}.130$		    \\ \hline
	\end{tabularx}
	\vspace{-0.2cm}
	\caption{Ablation study on the local image loss on Oxford-102. See text for detailed explanation.} \label{tab:ablation-loss} \vspace{-0.3cm}
\end{table}
\begin{figure}[t]
	\centering
	\includegraphics[width=0.491\textwidth,height=0.29\textwidth]{figure/imgloss.pdf}
	\vspace{-.6cm}
	\caption{Qualitative evaluation of the local image loss. The two images w/ the local image loss more precisely exhibit the complex flower petal structures described in the (colored) text.} \label{fig:vis-imgloss}
	\vspace{-.2cm}
\end{figure}

{\small
\bibliographystyle{ieee}
\bibliography{reference_zizhao,egbib}
}

\end{document}